Industrial, large-scale model predictive control with structured neural networks

### 2020-11-03 paper.tex
jbr edits on text

@@

% \biboptions{}
\newcommand{\useq}{\mathbf{u}}
\newcommand{\xseq}{\mathbf{x}}
\newcommand{\bbR}{\mathbb{R}}
\newcommand{\bbW}{\mathbb{W}}
\newcommand{\bbU}{\mathbb{U}}
\newcommand{\bbI}{\mathbb{I}}
\newcommand{\bbX}{\mathbb{X}}
\newcommand{\bbP}{\mathbb{P}}

@@

Model predictive control (MPC) is an online 
optimization based feedback control technology. A dynamic model 
of the plant is used to make forecasts of the plant measurements
in response to the actuator movements, and an optimization
is solved in real time to determine the optimal actuator
move to be applied to the plant. For linear plant models, 
the optimization to be solved online is a quadratic program (QP).
The development 
of efficient QP algorithms 
\citep*{kouzoupis:frison:zanelli:diehl:2018, wright:2019}
allowed practitioners to adopt MPC
in process industries \citep*{qin:badgwell:2003, lahiri:2017}.

An alternative strategy to 
real time optimization for the deployment of MPC 
is to characterize the MPC feedback law
as a piecewise affine function defined on 
polyhedral partitions of the state space, 
and use the feedback law online
\citep*{bemporad:morari:dua:pistikopoulos:2002, 
seron:goodwin:dedona:2003}.
The real time computation in this approach is 
reduced to a table look-up in the state space
based on the current value of parameters 
in the QP such as the state estimate and some 
steady-state targets that are computed for 
offset-free control. The number of polyhedral 
partitions of the state space grows exponentially
with model dimensions, however. Thereby prohibiting 
the deployment of this strategy on large-scale 
systems.

@@

To make progress in the size of control applications 
achievable with MPC, large problems
in which QP solvers fail to deliver the 
control in the available sample time should be addressed.
The scalability of the NN approach
to such problems with large model dimensions and 
control horizon length in the QP
has not been demonstrated. Two issues arise 
in these problems:
i) In a large dimensional state space,   
the entire set of feasible states of the QP cannot be sampled  
densely for the NN training. This issue is 
a form of the curse of dimensionality.
ii) One may sample the state space partially
to avoid this problem, but the time required for the 
data generation can still be impractical
due to the long time required to solve a QP 
for each sampled state. 
We demonstrate in this article by case studies that
this latter issue can be resolved for a large size 
of problems with parallel 
computing, and the former issue by sampling the 
state space for typical plant operational scenarios.
The following feature must be present in the large-scale
application of interest for the partial sampling scheme 
proposed in this article to be applicable:
for a particular operating mode of the plant, 
the number of frequently changing setpoints
and large magnitude disturbances 
must be reasonably small 
such that the entire large
dimensional state space is not visited during that
operation mode. 
The MPC feedback law 
can then be reliably approximated using NNs
for the different modes of plant operations driven by  
their respective set
of setpoint changes and disturbances.

From the above viewpoint, 
this article discusses the design of NNs
as an alternative to real time optimization in MPC, 
with an emphasis on large applications which 
are challenging with QP solvers. 
The focus of this article is on setpoint tracking 
MPC problems relevant for deployment in process industries.
We present a new feedforward network architecture to obtain
offset-free closed-loop performance with the NNs.
Additionally, we use the input-to-state
stability results \citep*{sontag:wang:1995}
available in the MPC literature, and establish
conditions under which the NN controller 
is robust to state estimation errors and process 
disturbances. We present case studies 
on process control examples which
demonstrate the scalability of the proposed NN design 
approach.

The related work by \cite*{chen:wang:atanasov:kumar:morari:2019}
demonstrates the scalability of NNs 
to approximate the MPC feedback law on systems with 
state dimensions upto thirty-six. The setpoint 
tracking offset-free MPC problem and state 
dimensions upto two-fifty considered in this 
article is not addressed in that work. 
\cite*{drgona:picard:kvasnica:helsen:2018}
takes a different approach for the scalability
of NN controllers, in which 
the dimensionality of the parameters in the MPC QP is first 
reduced using principal component analysis, and the 
NN controller is developed in a low dimensional state space.

@@

Notation. The symbols $\bbI$ and $\bbR$ are 
used to denote integers and reals respectively. 
Subscripts denote restrictions 
(e.g., $\bbR_{\geq 0}$ for nonnegative reals and 
$\bbI_{a:b}$ for integers in the 
closed interval $[a, b]$). The Euclidean norm 
is denoted by $\vert \cdot \vert$. A function 
$\alpha : \bbR_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{K}$ if it is continuous, zero 
at zero, and strictly increasing. This function is of class 
$\mathcal{K}_{\infty}$ if it is of class $\mathcal{K}$
and unbounded ($\alpha(s) \rightarrow \infty$ as 
$s \rightarrow \infty$). A function
$\beta : \bbR_{\geq 0} \times \bbI_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{K} \mathcal{L}$ if it is continuous,
for a fixed $k \in \bbI_{\geq 0}$, $\beta(\cdot, k)$ is of class $\mathcal{K}$,
and for a fixed $s \geq 0$, $\beta(s, \cdot)$ is nonincreasing
and satisfies $\textnormal{lim}_{k \rightarrow \infty} \beta(s, k) = 0$.
Given $V : X \rightarrow \bbR_{\geq 0}$ and $\tau > 0$, define 
$\textnormal{lev}_{\tau} V = \{ x \in X | V(x) \leq \tau \}$.
Bold symbols, e.g., $\mathbf{d}$ denote sequences, 
$d(k)$ denotes an element of $\mathbf{d}$ at time 
$k \in \bbI_{\geq 0}$, and $\mathbf{d}_i$ 
denotes the collection of elements of 
$\mathbf{d}$ for $k \in \bbI_{0:i-1}$.
Define $\vert \vert \mathbf{d}_i \vert \vert = 
\textnormal{max}_{k \in \bbI_{0:i-1}} \vert d(k) \vert$.
For a given vector $a$, lower and upper bounds 
($\underline{a}, \overline{a}$), define the saturation function 
as $\textnormal{sat}(a, \underline{a}, \overline{a}) = 
\{ a \ \textnormal{if} \ \underline{a} \leq a \leq \overline{a}; 
\underline{a} \ \textnormal{if} \ a < \underline{a};
\overline{a} \ \textnormal{if} \ \overline{a} < a \}$.

@@

\subsection{Model}
We consider a linear discrete time model augmented with an 
integrating disturbance model \citep{pannocchia:rawlings:2003}
\begin{align}\label{eq:ltimodel}
    x^+ = Ax + Bu + B_dd, \quad d^+ = d, \quad y = Cx + C_dd
\end{align}
in which $x \in \bbR^n$ is the state, 
$u \in \bbR^m$ is the control input, and
$d \in \bbR^d$ is the state of the disturbance
model. The matrices $A \in \bbR^{n \times n}$, 
$B \in \bbR^{n \times m}$,  
$C \in \bbR^{p \times n}$ denote the actuator to measurement 
model, and $B_d \in \bbR^{n \times d}$, $C_d \in \bbR^{p \times d}$
is the disturbance model.
The objective of the disturbance model
is to remove offset in the controlled 
plant measurements at steady
state operations and maintain
them at their setpoints.
We assume in the rest of this 
article that a Kalman filter can be constructed to 
estimate the model state ($\hat{x}$) and disturbance ($\hat{d}$) 
from measurements.

@@

\subsection{Regulator}
Based on the state estimate and
target steady state pair,
the following QP is solved to determine the 
actuator move to be applied to the plant
\begin{align} \label{eqn:regulator}
    V_N(\tilde{x}(0), \mathbf{\tilde{u}}) = \dfrac{1}{2}\sum_{k=0}^{N-1} & \Big(|\tilde{x}(k)|^2_Q  + |\tilde{u}(k)|^2_R \Big) + \dfrac{1}{2}|\tilde{x}(N)|^2_P
\end{align}
subject to 
\begin{align}
    \tilde{x}^+ &= A\tilde{x} + B\tilde{u}, \quad \tilde{x}(0) = \hat{x} - x_s \\
    &\underline{u} \leq \tilde{u} + u_s \leq \overline{u}
    \label{eqn:regulator_constraints}
\end{align}
in which $\hat{x}$ is the state estimate after the current measurement,
$N$ is the control 
horizon length, and
($\tilde{x}, \tilde{u}$) are the state and control 
in deviation from the current steady state targets ($x_s, u_s$). 
The penalty matrix $P$ is chosen as the 
optimal cost-to-go matrix of the unconstrained 
linear quadratic infinite horizon problem.
The decision variable in this 
QP is the control sequence 
$\mathbf{\tilde{u}} = [\tilde{u}(0)', \tilde{u}(1)', ..., \tilde{u}(N-1)']'$.
The first move of the optimal solution is denoted 
as the MPC feedback law,
$\kappa_N(\hat{x}, x_s, u_s) = 
\mathbf{\tilde{u}}^0(0;\hat{x}, x_s, u_s) + u_s$, 
which is applied to the plant. This feedback law is a function of the 
state estimate and target steady-state pair. A rate-of-change 
penalty on the control can be imposed in the QP objective
by augmenting the dynamic model with a surrogate state 
for the control as discussed in \cite*{rao:rawlings:1999}. 
In this case, 
the control injected at the previous timestep
($u_{-1}$) is supplied to the QP, 
and the feedback law becomes a function of the previous control 
as well in addition to the state estimate
and target steady-state pair.

@@

\subsection{Structured neural network}
An intuitive approach is to build a feedforward NN 
that takes the triple ($x, x_s, u_s$) as its input 
and outputs a control that is close to the optimal control. 
This strategy has been proposed in \cite*{karg:lucia:2020} and 
\cite*{chen:saulnier:atanasov:lee:kumar:pappas:morari:2018}
for one fixed steady state. The approach has not been 
demonstrated to scale to the size of model dimensions 
and the setpoint tracking 
offset-free MPC problem considered in this article.
An issue with this strategy 
is that the NN has no knowledge about 
the MPC feedback law, and can require large amounts
of training data in large state dimensions 
to obtain a reasonable controller. At a minimum,
the MPC feedback law is such that at the origin, the 
control action taken by the MPC controller is zero. For the 
setpoint tracking MPC problem considered in this article, 
the NN control law must be such that at steady-state 
operations, the NN maintains the plant at the desired steady state
such that the controlled measurements are at their setpoints.
This structure implies $u_s = \kappa_{NN}(x = x_s, x_s, u_s)$, 
in which we use $\kappa_{NN}(\cdot)$ to denote the NN control law.
To incorporate this information in a NN, we introduce the following
architecture

\begin{align*}
z_0 &= \begin{bmatrix}
    x', & x_s', & u_s', & x_s', & x_s', & u_s'
    \end{bmatrix}'\\
    f_i(z_{i-1}) &= \begin{bmatrix}
        W_i & 0 \\
        0 & W_i
          \end{bmatrix}z_{i-1}  + \begin{bmatrix}
            b_i \\
            b_i
            \end{bmatrix} \\
    z_i &= g(f_i(z_{i-1})), \textnormal{for} \ i \in \bbI_{1:h} \\
    u &= u_s + \begin{bmatrix}
  W_{h + 1}, & -W_{h + 1} \\
    \end{bmatrix} z_h \\
\end{align*}
in which $g(a) = \textnormal{max} (0, a)$,
is the ReLU operation applied element-wise on its vector input,
$z_i$ is the output of each hidden layer
of the network, $z_0$ is the input to the network, 
$h$ is the number of hidden layers, 
and $i \in \bbI_{1:h}$. 
The parameters to be optimized using 
training data are ($W_i, b_i$) 
and $W_{h+1}$, which are matrices 
of appropriate dimensions.
At steady-states
when $x=x_s$ in $z_0$, this structured network
outputs $u_s$ regardless of the choice of weights.
The software tensorflow 
\citep{abadi:agarwal:barham:brevdo:et-al:2015} is used 
for NN training purposes in this article,
and the structured architecture can be built 
with symbolic differentiation tools available in tensorflow.
When a rate-of-change penalty on the control 
is included in the regulator QP, 
the input to the structured NN is modified to 
\begin{align*}
    z_0 &= 
    \begin{bmatrix}
    x', & u_{-1}', & x_s', & u_s', & x_s', & u_s', & x_s', & u_s' \\
    \end{bmatrix}'
\end{align*}
in which $u_{-1}$ is the previous control applied 
to the plant and the dimensions of the matrices 
($W_i, b_i$) and $W_{h+1}$ 
are appropriately adjusted. Even after 
training, the control produced by 
the NN may not satisfy the hard actuator constraints,
and the saturation function 
($\textnormal{sat}(\cdot, \underline{u}, \overline{u})$) 
is applied to the output of the NN.

@@

For one fixed steady state, the 
set of feasible states for the MPC QP 
defined by \eqref{eqn:regulator}-\eqref{eqn:regulator_constraints}
is the entire $n$-dimensional state space $\bbR^n$.
Additionally, the target steady state may change
during the plant operation depending
on the setpoint and disturbance estimate. 
The domain of the MPC control law ($\kappa_N(x, x_s, u_s)$)
is therefore the state space 
$\bbR^{n}$ and the possible set of steady-state 
target pairs. Sampling this entire domain of the
MPC control law for NN training is difficult in larger 
dimensions and we show in Section 
\ref{sec:simulation_studies} that it is
not required for practical applications.

From a plant operations perspective,
a majority of the large-scale petrochemical plants
operate in a relatively small number of operating regimes
or scenarios. Each operational scenario is driven 
by a selected few controlled measurement 
setpoints that depend on product demands
and some large magnitude disturbances
may change frequently, 
while the setpoints for 
the other measurements remain constant for long periods 
of time. The states and steady-state targets 
spanned in a closed-loop 
operation per scenario is thereby considerably less 
than the entire domain of the MPC control law.
This type of operation of large chemical plants
can be exploited, and we propose the 
following data generation procedure 
to perform offline simulations using the model to 
collect the operationally 
relevant set of ($x, x_s, u_s, \kappa_N(\cdot)$) 
for the NN training:

@@

As reviewed in \cite*{mayne:rawlings:rao:scokaert:2000},
the nominal stability of MPC can be established
using the optimal cost function ($V_N^0(x)$) of the QP
as a Lyapunov function.
The region of attraction of the closed-loop system 
$x^+ = Ax + B\kappa_N(x)$ when no hard terminal region 
constraint \citep*{limon:alamo:salas:camacho:2006}  
is included in the QP can be characterized
as $\mathcal{X}_N = \textnormal{lev}_{Nd + \tau} \ V_N^0(x)$.
In which, $\tau$ is a level
set parameter of the terminal region chosen as
$\bbX_f = \textnormal{lev}_{\tau} \ x'Px$, and $d > 0$
is a constant such that $\ell(x, u) \geq d$ for all 
$x \in \bbR^n \setminus \bbX_f$ and 
$\underline{u} \leq u+u_s \leq \overline{u}$.
The parameter $\tau$ is chosen small enough such that
the unconstrained LQR satisfies the actuator constraints 
for all $x \in \bbX_f$. The optimal cost 
function is continuous and satisfies 
$c_1\vert x \vert^2 \leq V_N^0(x) \leq c_2\vert x \vert^2$
and $V_N^0(x^+) - V_N^0(x) \leq -c_1\vert x \vert^2$ 
for some $c_2 \geq c_1 > 0$.

Assume that a state estimate ($\hat{x}$) is
used by the NN to compute the control. 
To analyze robustness, the following perturbed
linear system is considered
\begin{align} \label{eq:perturbed_sys}
    \hat{x}^+ = A\hat{x} + B\kappa_N(\hat{x}) + Be_{NN}(\hat{x}) + w - Ae + e^+
\end{align}
in which $\hat{x} = x + e$ is the state estimate, 
$w$ is a process disturbance, 
and $e$, $e^+$ are the state estimation errors at the 
current and next timestep respectively. 
We use $\phi_d(k;\hat{x})$ to denote a 
solution of the perturbed system 
at time $k$ starting from an initial state $\hat{x}$,
and $\phi(k;x)$ to denote a solution 
of the nominal system when the state estimation errors and process
disturbances are zero starting 
from an initial state $x$.

@@

The proof of this theorem along with the required 
input-to-state stability definitions is provided 
in Appendix \ref{app:theorem1}.
In the nominal case, 
the disturbance sequences $\mathbf{e}$ and $\mathbf{w}$ are zero.
The closed-loop states in this case have the bound
$\vert \phi(k; x) \vert \leq \beta(\vert x \vert, k) + 
\alpha_n(\bar{e}_{NN})$. Any additional value of 
the NN training is to reduce the worst 
approximation error ($\bar{e}_{NN}$) 
in the state space of interest and 
strengthen the bounds on the closed-loop states.

@@

In this section, two case studies are presented 
which demonstrate the 
scalability of the proposed NN design approach.
The closed-loop performance of NNs and computational benefits over 
online QP based MPC are analyzed in these case studies.
After the offline design of NNs,
two types of validation simulations 
are performed: (i) with setpoints 
and disturbances not present in the training 
data but within the same range of values used to 
generate the training data, (ii) with extra setpoint 
signals assumed to be
unknown during the offline network design phase.
Both these types of simulations are conducted to 
analyze the interpolation and extrapolation 
abilities of the NNs, and are referred subsequently as 
simulations with expected and unexpected plant behaviors.

The closed-loop performance of the 
following alternative and 
equivalently fast controllers as NNs are also compared
in the validation simulations.
a) steady state controller (SS): the solution
of the target selector is directly applied to the plant 
$u = u_s$, b) saturated linear quadratic regulator (satK):
the unconstrained LQR gain 
$K$ is computed and used in the control law 
$u = \textnormal{sat}(K(x-x_s) + u_s, \underline{u}, \overline{u})$,
c) short horizon controller (SH): an MPC problem
with a short control horizon is solved online, 
d) Unstructured NN (NN-UNS): a standard feedforward network is 
trained with the triple ($x, x_s, u_s$) as the input
and then used as the feedback controller online.

@@

\section{Conclusion} \label{sec:conclusion}
Neural networks have been demonstrated 
to approximate the MPC feedback law
for treating large-scale, linear MPC applications
that may be out of reach with QP solvers. 
The proposed no online optimization based
methods to date in the literature such as storing 
the entire MPC feedback law 
and sampling the large dimensional state space 
for the NN training are not tractable for the problems 
considered in this article.

>>>>

\usepackage{mpcsymbols}

@@

% \biboptions{}

@@

Model predictive control (MPC) is an online 
optimization based feedback control technology. A dynamic model 
of the plant is used to make forecasts of the plant measurements
in response to the actuator movements, and an optimization
is solved in real time to determine the optimal actuator
move to be applied to the plant. For linear plant models, 
the optimization to be solved online is a quadratic program (QP).
The development 
of efficient QP algorithms 
\citep*{kouzoupis:frison:zanelli:diehl:2018, wright:2019}
allowed practitioners to apply MPC
in the process industries \citep*{qin:badgwell:2003, lahiri:2017}.

An alternative strategy to 
real time optimization for the deployment of MPC 
is to characterize the MPC feedback law
as a piecewise affine function defined on 
polyhedral partitions of the state space, 
and use the feedback law online
\citep*{bemporad:morari:dua:pistikopoulos:2002, 
seron:goodwin:dedona:2003}.
The real time computation in this approach is 
reduced to a table look-up in the state space
based on the current value of parameters 
in the QP such as the state estimate and some 
steady-state targets that are computed for 
offset-free control. The number of polyhedral 
partitions of the state space grows exponentially
with model dimensions, however, which prohibits
the deployment of this strategy on large-scale 
systems.

@@

To make progress in the size of control applications 
achievable with MPC, large problems
in which QP solvers fail to deliver the 
control in the available sample time should be addressed.
The scalability of the NN approach
to such problems with large model dimensions and 
control horizon length in the QP
has not been demonstrated. Two issues arise 
in these problems:
i) In a large dimensional state space,   
the entire set of feasible states of the QP cannot be sampled  
densely for the NN training. This issue is 
a form of the well-known curse of dimensionality.
ii) One may sample the state space partially
to avoid this problem, but the time required for the 
data generation can still be impractical
due to the long time required to solve a QP 
for each sampled state. 
We demonstrate in this article by case studies that
this latter issue can be resolved for a large size 
of problems with parallel 
computing, and the former issue by sampling the 
state space for typical plant operational scenarios.
The following feature must be present in the large-scale
application of interest for the partial sampling scheme 
proposed in this article to be applicable:
for a particular operating mode of the plant, 
the number of frequently changing setpoints
and large magnitude disturbances 
must be reasonably small 
such that the entire large
dimensional state space is not visited during that
operation mode. 
The MPC feedback law 
can then be reliably approximated using NNs
for the different modes of plant operations driven by  
their respective set
of setpoint changes and disturbances.

From the above viewpoint, 
this article discusses the design of NNs
as an alternative to real time optimization in MPC, 
with an emphasis on large applications which 
are challenging with QP solvers. 
The focus of this article is on setpoint tracking 
MPC problems relevant for deployment in the process industries.
We present a new feedforward network architecture to obtain
offset-free closed-loop performance with the NNs.
Additionally, we use the input-to-state
stability results
available in the MPC literature, and establish
conditions under which the NN controller 
is robust to state estimation errors and process 
disturbances. We present case studies 
on process control examples that
demonstrate the scalability of the proposed NN design 
approach.

The related work by \cite*{chen:wang:atanasov:kumar:morari:2019}
demonstrates the scalability of NNs 
to approximate the MPC feedback law on systems with 
state dimensions up to 36. The setpoint 
tracking offset-free MPC problem and state 
dimensions up to 250 considered in this 
article is not addressed in that work. 
\cite*{drgona:picard:kvasnica:helsen:2018}
take a different approach for the scalability
of NN controllers, in which 
the dimensionality of the parameters in the MPC QP is first 
reduced using principal component analysis, and the 
NN controller is developed in a low dimensional state space.

@@

Notation. The symbols $\bbI$ and $\bbR$ are 
used to denote integers and reals respectively. 
Subscripts denote restrictions 
(e.g., $\bbR_{\geq 0}$ for nonnegative reals and 
$\bbI_{a:b}$ for integers in the 
closed interval $[a, b]$). The Euclidean norm 
is denoted by $\vert \cdot \vert$. A function 
$\alpha : \bbR_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{K}$ if it is continuous, zero 
at zero, and strictly increasing. This function is of class 
$\mathcal{K}_{\infty}$ if it is of class $\mathcal{K}$
and unbounded ($\alpha(s) \rightarrow \infty$ as 
$s \rightarrow \infty$). A function
$\beta : \bbR_{\geq 0} \times \bbI_{\geq 0} \rightarrow \bbR_{\geq 0}$
is of class $\mathcal{KL}$ if it is continuous,
for each $k \in \bbI_{\geq 0}$, $\beta(\cdot, k)$ is of class $\mathcal{K}$,
and for each $s \geq 0$, $\beta(s, \cdot)$ is nonincreasing
and satisfies $\textnormal{lim}_{k \rightarrow \infty} \beta(s, k) = 0$.
Given $V : X \rightarrow \bbR_{\geq 0}$ and $\tau > 0$, define 
$\textnormal{lev}_{\tau} V = \{ x \in X \mid V(x) \leq \tau \}$.
Bold symbols, e.g., $\mathbf{d}$ denote sequences, 
$d(k)$ denotes an element of $\mathbf{d}$ at time 
$k \in \bbI_{\geq 0}$, and $\mathbf{d}_i$ 
denotes the collection of elements of 
$\mathbf{d}$ for $k \in \bbI_{0:i-1}$.
Define $\vert \vert \mathbf{d}_i \vert \vert = 
\textnormal{max}_{k \in \bbI_{0:i-1}} \vert d(k) \vert$.
For a given vector $a$, lower and upper bounds 
($\underline{a}, \overline{a}$), define the saturation function 
as $\textnormal{sat}(a, \underline{a}, \overline{a}) = 
\{ a \ \textnormal{if} \ \underline{a} \leq a \leq \overline{a}; 
\underline{a} \ \textnormal{if} \ a < \underline{a};
\overline{a} \ \textnormal{if} \ \overline{a} < a \}$.

@@

\subsection{Model}
We consider a linear discrete time model augmented with an 
integrating disturbance model \citep{pannocchia:rawlings:2003}
\begin{align}\label{eq:ltimodel}
    x^+ = Ax + Bu + B_dd, \quad d^+ = d, \quad y = Cx + C_dd
\end{align}
in which $x \in \bbR^n$ is the state, 
$u \in \bbR^m$ is the control input, and
$d \in \bbR^d$ is the state of the disturbance
model. The matrices $A \in \bbR^{n \times n}$, 
$B \in \bbR^{n \times m}$,  
$C \in \bbR^{p \times n}$ denote the actuator to measurement 
model, and $B_d \in \bbR^{n \times d}$, $C_d \in \bbR^{p \times d}$
is the disturbance model.
The objective of the disturbance model
is to remove offset in the controlled 
plant measurements at steady
state operations and maintain
the controlled variables at their setpoints.
We assume in the rest of this 
article that a Kalman filter can be constructed to 
estimate the model state ($\hat{x}$) and disturbance ($\hat{d}$) 
from measurements.

@@

\subsection{Regulator}
Based on the state estimate and
target steady state pair,
the following QP is solved to determine the 
actuator move to be applied to the plant
\begin{align} \label{eqn:regulator}
    V_N(\tilde{x}(0), \mathbf{\tilde{u}}) = \dfrac{1}{2}\sum_{k=0}^{N-1} & \Big(|\tilde{x}(k)|^2_Q  + |\tilde{u}(k)|^2_R \Big) + \dfrac{1}{2}|\tilde{x}(N)|^2_P
\end{align}
subject to 
\begin{align}
    \tilde{x}^+ &= A\tilde{x} + B\tilde{u}, \quad \tilde{x}(0) = \hat{x} - x_s \\
    &\underline{u} \leq \tilde{u} + u_s \leq \overline{u}
    \label{eqn:regulator_constraints}
\end{align}
in which $\hat{x}$ is the state estimate after the current measurement,
$N$ is the control 
horizon length, and
($\tilde{x}, \tilde{u}$) are the state and control 
in deviation from the current steady state targets ($x_s, u_s$). 
The penalty matrix $P$ is chosen as the 
optimal cost-to-go matrix of the unconstrained infinite horizon 
linear quadratic problem.
The decision variable in this 
QP is the control sequence 
$\mathbf{\tilde{u}} = [\tilde{u}(0)', \tilde{u}(1)', ..., \tilde{u}(N-1)']'$.
The first move of the optimal solution is denoted 
as the MPC feedback law,
$\kappa_N(\hat{x}, x_s, u_s) = 
\mathbf{\tilde{u}}^0(0;\hat{x}, x_s, u_s) + u_s$, 
which is applied to the plant. This feedback law is a function of the 
state estimate and target steady-state pair. A rate-of-change 
penalty on the control can be imposed in the QP objective
by augmenting the dynamic model with a surrogate state 
for the control as discussed in \cite*{rao:rawlings:1999}. 
In this case, 
the control injected at the previous timestep
($u_{-1}$) is supplied to the QP, 
and the feedback law becomes a function of the previous control 
as well in addition to the state estimate
and target steady-state pair.

@@

\subsection{Structured neural network}
An intuitive approach is to build a feedforward NN 
that takes the triple ($x, x_s, u_s$) as its input 
and outputs a control that is close to the optimal control. 
This strategy has been proposed in \cite*{karg:lucia:2020} and 
\cite*{chen:saulnier:atanasov:lee:kumar:pappas:morari:2018}
for a single fixed steady state. The approach has not been 
demonstrated to scale to the size of model dimensions 
and the setpoint tracking 
offset-free MPC problem considered in this article.
An issue with this strategy 
is that the NN has no knowledge about 
the MPC feedback law, and can require large amounts
of training data in large state dimensions 
to obtain a reasonable controller. At a minimum,
the MPC feedback law is such that at the origin, the 
control action taken by the MPC controller is zero. For the 
setpoint tracking MPC problem considered in this article, 
the NN control law must be such that at steady-state 
operations, the NN maintains the plant at the desired steady state
such that the controlled measurements are at their setpoints.
This structure implies $u_s = \kappa_{NN}(x = x_s, x_s, u_s)$, 
in which we use $\kappa_{NN}(\cdot)$ to denote the NN control law.
To incorporate this information in a NN, we introduce the following
architecture
\begin{align}
z_0 &= \begin{bmatrix}
    x', & x_s', & u_s', & x_s', & x_s', & u_s' \end{bmatrix}'\notag \\
    f_i(z_{i-1}) &= \begin{bmatrix}
        W_i & 0 \notag \\
        0 & W_i
          \end{bmatrix}z_{i-1}  + \begin{bmatrix}
            b_i \notag 
            b_i
            \end{bmatrix} \notag\\
    z_i &= g(f_i(z_{i-1})), \textnormal{for} \ i \in \bbI_{1:h} \notag \\
    u &= u_s + \begin{bmatrix}
  W_{h + 1}, & -W_{h + 1} \\
    \end{bmatrix} z_h
\label{eq:NNstruct}
\end{align}
in which $g(a) = \textnormal{max} (0, a)$,
is the ReLU operation applied element-wise on its vector input,
$z_i$ is the output of each hidden layer
of the network, $z_0$ is the input to the network, 
$h$ is the number of hidden layers, 
and $i \in \bbI_{1:h}$. 
The parameters to be optimized using 
training data are ($W_i, b_i$) 
and $W_{h+1}$, which are matrices 
of appropriate dimensions.
At steady state, when $x=x_s$ in $z_0$, \eqref{eq:NNstruct}
outputs $u_s$ regardless of the choice of weights.
The software tensorflow 
\citep{abadi:agarwal:barham:brevdo:et-al:2015} is used 
for NN training purposes in this article,
and the structured architecture can be built 
with symbolic differentiation tools available in tensorflow.
When a rate-of-change penalty on the control 
is included in the regulator QP, 
the input to the structured NN is modified to 
\begin{align*}
    z_0 &= 
    \begin{bmatrix}
    x', & u_{-1}', & x_s', & u_s', & x_s', & u_s', & x_s', & u_s' \\
    \end{bmatrix}'
\end{align*}
in which $u_{-1}$ is the previous control applied 
to the plant and the dimensions of the matrices 
($W_i, b_i$) and $W_{h+1}$ 
are appropriately adjusted. Even after 
training, the control produced by 
the NN may not satisfy the hard actuator constraints,
and the saturation function 
($\textnormal{sat}(\cdot, \underline{u}, \overline{u})$) 
is applied to the output of the NN.

@@

For one fixed steady state, the 
set of feasible states for the MPC QP 
defined by \eqref{eqn:regulator}-\eqref{eqn:regulator_constraints}
is the entire $n$-dimensional state space $\bbR^n$.
Additionally, the target steady state may change
during the plant operation depending
on the setpoint and disturbance estimate. 
The domain of the MPC control law ($\kappa_N(x, x_s, u_s)$)
is therefore the state space 
$\bbR^{n}$ and the possible set of steady-state 
target pairs. Sampling this entire domain of the
MPC control law for NN training is difficult in large 
dimension, and we show in Section 
\ref{sec:simulation_studies} that it is
also not required for practical applications.

From a plant operations perspective,
a majority of the large-scale petrochemical plants
operate in a relatively small number of operating regimes
or scenarios. Each operational scenario is driven 
by a selected few controlled measurement 
setpoints that depend on product demands
and some large magnitude disturbances
may change frequently, 
while the setpoints for 
the other measurements remain constant for long periods 
of time. The states and steady-state targets 
spanned in a closed-loop 
operation per scenario is thereby considerably less 
than the entire domain of the MPC control law.
This typical operation of large chemical plants
can be exploited, and we propose the 
following data generation procedure 
to perform offline simulations using the model to 
collect the operationally 
relevant set of ($x, x_s, u_s, \kappa_N(\cdot)$) 
for the NN training:

@@

As reviewed in \cite*{mayne:rawlings:rao:scokaert:2000},
the nominal stability of MPC can be established
using the optimal cost function ($V_N^0(x)$) of the QP
as a Lyapunov function.
The region of attraction of the closed-loop system 
$x^+ = Ax + B\kappa_N(x)$ when no hard terminal region 
constraint is included in the QP
can be characterized as $\mathcal{X}_N = \textnormal{lev}_{Nd + \tau} \
V_N^0(x)$ \citep*{limon:alamo:salas:camacho:2006}.
The parameter $\tau$ is the level set
parameter of the terminal region chosen as
$\bbX_f = \textnormal{lev}_{\tau} \ x'Px$, and $d > 0$
is a constant such that $\ell(x, u) \geq d$ for all 
$x \in \bbR^n \setminus \bbX_f$ and 
$\underline{u} \leq u+u_s \leq \overline{u}$.
The parameter $\tau$ is chosen small enough such that
the unconstrained LQR satisfies the actuator constraints 
for all $x \in \bbX_f$. The optimal cost 
function is continuous and satisfies 
$c_1\vert x \vert^2 \leq V_N^0(x) \leq c_2\vert x \vert^2$
and $V_N^0(x^+) - V_N^0(x) \leq -c_1\vert x \vert^2$ 
for some $c_2 \geq c_1 > 0$.

Assume that a state estimate ($\hat{x}$) is
used by the NN to compute the control. 
To analyze robustness, the following perturbed
linear system is considered
\begin{align} \label{eq:perturbed_sys}
    \hat{x}^+ = A\hat{x} + B\kappa_N(\hat{x}) + Be_{NN}(\hat{x}) + w - Ae + e^+
\end{align}
in which $\hat{x} = x + e$ is the state estimate, 
$w$ is a process disturbance, 
and $e$, $e^+$ are the state estimation errors at the 
current and next timestep respectively. 
We use $\phi_d(k;\hat{x})$ to denote a solution of the perturbed system  at time
$k$ starting from an initial state $\hat{x}$; and $\phi(k;x)$ to denote a
solution  of the nominal system starting  from an initial state $x$ when the
state estimation errors and process disturbances are zero.

@@

The proof of this theorem along with the required 
input-to-state stability definitions are provided 
in Appendix \ref{app:theorem1}.
In the nominal case, 
the disturbance sequences $\mathbf{e}$ and $\mathbf{w}$ are zero.
The closed-loop states in this case have the bound
$\vert \phi(k; x) \vert \leq \beta(\vert x \vert, k) + 
\alpha_n(\bar{e}_{NN})$. Any additional value of 
the NN training is to reduce the worst 
approximation error ($\bar{e}_{NN}$) 
in the state space of interest and 
reduce the bound on the closed-loop states.

@@

In this section, two case studies are presented 
that demonstrate the 
scalability of the proposed NN design approach.
The closed-loop performance of NNs and computational benefits over 
online QP based MPC are analyzed in these case studies.
After the offline design of NNs,
two types of validation simulations 
are performed: (i) with setpoints 
and disturbances not present in the training 
data but within the same range of values used to 
generate the training data, (ii) with extra setpoint 
signals assumed to be
unknown during the offline network design phase.
Both these types of simulations are conducted to 
analyze the interpolation and extrapolation 
abilities of the NNs, and are referred subsequently as 
simulations with \textit{expected} and \textit{unexpected} plant behaviors, respectively.

The closed-loop performance of the 
following alternative and 
equivalently fast controllers as NNs are also compared
in the validation simulations:
(a) steady state controller (SS): the solution
of the target selector is directly applied to the plant 
$u = u_s$; (b) saturated linear quadratic regulator (satK):
the unconstrained LQR gain 
$K$ is computed and used in the control law 
$u = \textnormal{sat}(K(x-x_s) + u_s, \underline{u}, \overline{u})$;
(c) short horizon controller (SH): an MPC problem
with a short control horizon is solved online;
(d) Unstructured NN (NN-UNS): a standard feedforward network is 
trained with the triple ($x, x_s, u_s$) as the input
and then used as the feedback controller online.

@@

\section{Conclusion} \label{sec:conclusion}
Neural networks have been demonstrated 
to approximate the MPC feedback law
for treating large-scale, linear MPC applications
that may be out of reach with QP solvers. 
The proposed methods avoiding online optimization
to date in the literature, such as storing 
the entire MPC feedback law 
and sampling the large dimensional state space 
for the NN training, are not tractable for the problems 
considered in this article.
